---
title: "Lab 9. Linear mixed-effect models"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## [Assignment link](https://docs.google.com/forms/d/e/1FAIpQLSfQAI87CVsjqbkw1XbvJw_dmj8tET1s-oT9jmQGx9miRRtcMQ)

```{r, message= FALSE, include=TRUE}
library(tidyverse)
#install.packages("lingtypology")
library(lingtypology)
library(lme4)
```

### 1. Consonants and vowels
```{r, warning=FALSE, include=TRUE}
lapsyd <- read.csv("https://goo.gl/eD4S5n")
map.feature(lapsyd$name, features = lapsyd$area, 
              label = lapsyd$name,
              label.hide = TRUE)
```

```{r}
fit1 <- lm(count_vowel~count_consonant, data = lapsyd)
summary(fit1)

lapsyd$model1 <- predict(fit1)

lapsyd %>% 
  ggplot(aes(count_consonant, count_vowel))+
  geom_point()+
  geom_line(aes(count_consonant, model1), color = "blue")+
  labs(x = "number of consonants",
       y = "number of vowels",
       caption = "data from LAPSyD")

fit2 <- lmer(count_vowel ~ count_consonant + (1|area), data = lapsyd)
summary(fit2)

lapsyd$model2 <- predict(fit2)

lapsyd %>% 
  ggplot(aes(count_consonant, count_vowel))+
  geom_point()+
  geom_line(aes(count_consonant, model2, color = area))+
  labs(x = "number of consonants",
       y = "number of vowels",
       caption = "data from LAPSyD")

fit3 <- lmer(count_vowel ~ count_consonant + (1+count_consonant|area), data = lapsyd)
summary(fit3)

lapsyd$model3 <- predict(fit3)

lapsyd %>% 
  ggplot(aes(count_consonant, count_vowel))+
  geom_point()+
  geom_line(aes(count_consonant, model3, color = area))+
  labs(x = "number of consonants",
       y = "number of vowels",
       caption = "data from LAPSyD")

fit4 <- lmer(count_vowel ~ count_consonant + (0+count_consonant|area), data = lapsyd)
summary(fit4)

lapsyd$model4 <- predict(fit4)

lapsyd %>% 
  ggplot(aes(count_consonant, count_vowel))+
  geom_point()+
  geom_line(aes(count_consonant, model4, color = area))+
  labs(x = "number of consonants",
       y = "number of vowels",
       caption = "data from LAPSyD")

fit5 <- lmer(count_vowel ~ count_consonant + (1|area) + (0+count_consonant|area), data = lapsyd)
summary(fit5)

lapsyd$model5 <- predict(fit5)

lapsyd %>% 
  ggplot(aes(count_consonant, count_vowel))+
  geom_point()+
  geom_line(aes(count_consonant, model5, color = area))+
  labs(x = "number of consonants",
       y = "number of vowels",
       caption = "data from LAPSyD")
```

```{r}
anova(fit5, fit4, fit3, fit2, fit1)
```

## 2. Icelandic vowels
This set is based on (Coretta 2017, https://goo.gl/NrfgJm). This dissertation deals with the relation between vowel duration and aspiration in consonants. Author carried out a data collection with 5 natives speakers of Icelandic. Then he extracted the duration of vowels followed by aspirated versus non-aspirated consonants. Check out whether the vowels before consonants of different places of articulation are significantly different.

Use read.csv("https://goo.gl/7gIjvK") for downloading data.

### 2.1
Calculate mean values for vowel duration in your data grouped by `place` (of articulation) and `speaker`.

```{r}
df <- read.csv("https://goo.gl/7gIjvK")

df %>% 
  group_by(place, speaker) %>% 
  summarise(mean(vowel.dur))
# 
```
Let's do some visualization

```{r}
df %>% 
  ggplot(aes(place, vowel.dur)) + 
    geom_point(alpha = 0.2) + 
    facet_wrap(~ speaker) + 
    xlab("place of articulation") + 
    ylab("vowel duration")
```

### 2.2 Calculate mean values for vowel duration in your data grouped by `word`.
```{r}
df <- read.csv("https://goo.gl/7gIjvK")

df %>% 
  group_by(word) %>% 
  summarise(mean(vowel.dur))
# 
```

### 2.3 Fit mixed-effect linear regression model 
taking into account `speaker` as a random effect. Plot 

```{r}

fit <- lmer(vowel.dur ~ place + (1|speaker), data = df)
summary(fit)
plot(fit)
qqnorm(resid(fit))
qqline(resid(fit))
```

### 2.3 Fit mixed-effect linear regression model 
taking into account `speaker` and `word` as random effects. Note that random factors can be nested.


If our groups are nested (as it happens with speakers and words), the following model should be wrong:

```{r}
fit2.WRONG <- lmer(vowel.dur ~ place + (1|speaker) + (1|word), data = df)  # treats the two random effects as if they are crossed
```

To avoid future confusion let us create a new variable that is explicitly nested. Let???s call it sample:

```{r}
df <- within(df, sample <- factor(speaker:word))
head(summary(df$sample))
```

Now let's fit the nested mixed-effect model properly:
```{r}
fit2 <- lmer(vowel.dur ~ place + (1|speaker) + (1|sample), data = df)  # treats the two random effects as if they are nested
summary(fit2)
plot(fit2)
qqnorm(resid(fit2))
qqline(resid(fit2))

```
