---
title: "Lab14. CA and MCA."
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r 0.0, message=FALSE, warning=FALSE}
library(FactoMineR)
library(tidyverse)
library(ca)
library(vcd)
theme_set(theme_bw())
```

## Grammatical profiles of Russian verbs    

In their article "Predicting Russian aspect by frequency across genres" Eckhoff et al. (2017) ask whether the aspect of individual verbs can be predicted based on the statistical distribution of their inflectional forms. The dataset contains a sample of sentences taken from the Russian National Corpus. Each verb was annotated by Mood&Tense, Voice, Aspect and other grammatical features.    

```{r 1.01}
ru <- read.csv('https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/RNCverbSamples_journ.csv', sep=';')
str(ru)
```

First, we will do some preprocessing. Here, we join future passives with other passive participles.
```{r 1.02}
ru$MoodTense[ru$Voice == 'pass' & ru$MoodTense == 'indicfut'] <- "partcppast"
```

Let's look at top-10 of verbs in the dataset
```{r 1.03}
ru %>% 
  group_by(LemmaTranslit) %>% 
  summarize(count=n()) %>%
  arrange(desc(count))
```

### 1.1 Grammatical profiles
Grammatical profile is a vector that contains the number (or the ratio) of inflectional forms for individual lemmas.
We can pick up a subset containing only forms of the verb _chitat'_ 'read'.
```{r 1.1.1}
ru.chit <- droplevels(subset(ru, LemmaTranslit == "chitat'"))
```

This is the grammatical profile of _chitat'_
```{r 1.1.2}
print("The grammatical profile of chitat'")
print(table(ru.chit$MoodTense))
print(prop.table(table(ru.chit$MoodTense))*100)
```

Now we will calculate the grammatical profile of each verb (which has more than 50 occurrences in our data set) and split the resulting table into two parts: grammatical forms themselves (numeric data, see ttdata below) and metadata (categories labeled in the RNC or by annotators: lemma, trasitivity, aspect).

Table of tense-mood distribution per lemma:
```{r 1.1.3}
tab = table(ru$LemmaTranslit,ru$MoodTense)
#turns the table into a data frame
t = as.data.frame.matrix(tab)
#adds lemmas as a separate column
t$LemmaTranslit = row.names(t)
#adds metadata columns for transitivity and aspect, assuming that these are stable per lemma - just picking the first value per lemma
t$Trans = as.factor(unlist(lapply(t$LemmaTranslit, function(x) names(table(droplevels(subset(ru, LemmaTranslit == x)$Trans)))[1])))
t$Asp = as.factor(unlist(lapply(t$LemmaTranslit, function(x) names(table(droplevels(subset(ru, LemmaTranslit == x)$Aspect)))[1])))
```

Label the biaspectuals 'b':
```{r 1.1.4}
levels(t$Asp) <- c('i','p','b')
t[t$LemmaTranslit=="ispol'zovat'",]$Asp <- 'b'
t[t$LemmaTranslit=="obeschat'",]$Asp <- 'b'
```

Pick out the lemmas with 50 or more occurrences and split the data:
```{r 1.1.5}
tt <- t[rowSums(t[,1:9]) >= 50,]
```

## 1.2 t-SNE visualization (for numeric variables)   

The idea behind this kind of visualisation is to plot different clusters as far from each other as possible (preserving the distance between each pair of clusters). Within each cluster, the points are distributed to show the internal structure of the cluster. Note that unlike PCA, in t-SNE the points' coordinates cannot be interpreted directly (there is no linear mapping of one plane to another), and linear correlations can be misleading.  

```{r 1.2}
library(Rtsne)
ru.tsne <- Rtsne(tt[,1:9],
                 dims=2, 
                 perplexity=50, 
                 verbose=TRUE, 
                 max_iter = 2000)

tt <- cbind(tt, ru.tsne$Y)

tt %>% 
  ggplot(aes(`1`, `2`, label = Asp, color = Asp))+
  geom_text()
```

## CA 
```{r}
tt_ca <- ca(tt[,1:9])

tt_col <- data.frame(tt_ca$colcoord)
tt_col$rows <- rownames(tt_ca$colcoord)

tt_row <- data.frame(tt_ca$rowcoord)
tt_row$rows <- rownames(tt_ca$rowcoord)

tt_col %>% 
  ggplot(aes(Dim1, Dim2, label = rows))+
  geom_text(color = "blue")+
  geom_text(data = tt_row, color = "red")+
  labs(x = "Dim1 (38.63%)",
       y = "Dim2 (20.01%)")
```

## MCA 

We use `MCA()` function in `FactoMineR` to perform multiple correspondence analysis.

## 2 Grammaticalization of participles in Russian    

Kosheleva 2017 studied the adjectivation of participles in Russian as a gradient phenomenon. The participle form loses its verb-like properties when it is used in a particular context or is a form of a particular lexeme. In each sentence, the participle was evaluated as being more 'adjective-like' or more 'participle-like':  
* `Like`: adj vs. ptcp  
In addition, other grammatical properties of the participle forms were tagged:  
* `Tense`: past, pres(ent)  
* `Voice`: act(ive), pass(ive), med(medium, in reflexive verbs)  
* `Asp` (aspect): ipf (imperfective), pf (perfective)  
* `Trans` (transitivity): intr(ansitive), tran(sitive)  
* `Plen`: plen (full form), sh (short form)   
as well as the properties of construction they are used in:   
* `Cx`: attr(ibutive), depict(ive), pred(icative)  
* `Dep`: yes (the participle has dependencies), no  
* `Anim` (animacy of the subject): anim(ate), inan(imate)    

Load the data and take a peek
```{r 2.0}
adj <- read.csv(file='/Users/olyashevskaya/Documents/!Vyshka/DataAnalysis/2018/adj_ptcp_kosheleva.csv', sep='\t')
str(adj) # please check that all variables are factors. If not, apply as.factor()
summary(adj)
head(adj)
```

### 2.1 Number of levels per variable, descriptive statistics
We will use `cats` to color grammatical and constructional features in MCA plots:
```{r 2.1}
# apply MCA
cats = apply(adj, 2, function(x) nlevels(as.factor(x)))
cats
```

Contingency table with ggplot2:
```{r 2.1.1}
adj %>%
   group_by(Like, Voice) %>%
   summarise(number = n()) %>%
   ggplot(aes(Like, Voice, label = number))+
   geom_point(aes(size = number), color = "light blue")+
   geom_text()+
   scale_size(range = c(10, 30))+
   guides(size = F)+
   xlab("Adjective vs. Participle")+
   ylab("Voice")+
   ggtitle("Correlation between the level of adjectiveness and voice")
 
```

Mosaic plot, colored by Pearson residuals:
```{r 2.1.2}
mosaic(~ Like + Voice + Tense, data = adj, shade = T) # observed
```

We can also compare it with plot based on expected values:
```{r 2.1.3}
mosaic(~ Like + Voice + Tense, data = adj, type = "expected") # expected
```

### 2.2 MCA() 

```{r 2.2}
# apply MCA
mca1 = MCA(adj, graph = FALSE)
mca1
```

### 2.3 Inertia 
Table of eigenvalues will show us explained variance (inertia):
```{r 2.3}
mca1$eig
```

### 2.4 Plot
FactoMineR has a built-in function to plot the results:
```{r 2.4}
plot(mca1)
```

### 2.5 Plot variables with ggplot2
However, a nicer graph can be plotted with ggplot2
```{r 2.5}
# variable coordinates
mca1_vars_df = data.frame(mca1$var$coord, Variable = rep(names(cats), cats))

# individual observation coordinates
mca1_obs_df = data.frame(mca1$ind$coord)

# plot of variable categories
ggplot(data=mca1_vars_df, 
       aes(x = Dim.1, y = Dim.2, label = rownames(mca1_vars_df))) +
 geom_hline(yintercept = 0, colour = "gray70") +
 geom_vline(xintercept = 0, colour = "gray70") +
 geom_text(aes(colour=Variable)) +
 ggtitle("MCA plot of variables using FactoMineR and ggplot2")
```

### 2.6 Plot individuals and variables with ggplot2
```{r 2.6}
ggplot(data = mca1_obs_df, aes(x = Dim.1, y = Dim.2)) +
  geom_hline(yintercept = 0, colour = "gray70") +
  geom_vline(xintercept = 0, colour = "gray70") +
  geom_point(colour = "gray50", alpha = 0.7) +
  geom_density2d(colour = "gray60", alpha = 0.5) +
  geom_text(data = mca1_vars_df, 
            aes(x = Dim.1, y = Dim.2, 
                label = rownames(mca1_vars_df), colour = Variable)) +
  ggtitle("MCA plot of observations and variables using R package FactoMineR") +
  scale_colour_discrete(name = "Variable")
```

### 2.7 mjca()
This is another function from the package "ca" by Michael Greenacre and Oleg Nenadic.
Method `lambda="Burt"` is less conservative, so explained inertia can be higher:
```{r 2.7}
# apply mjca
mca2 <-  mjca(adj, lambda = "Burt", nd = 5)

#summary
# summary(mca2)

# eigenvalues
# mca2$sv^2

# column coordinates
head(mca2$colcoord, 3)

# row coordinates
head(mca2$rowcoord, 3)
```

### 2.8 Plot variables with ggplot2
Use the column coordinates `colcoord` to make a data frame and pass it to ggplot():
```{r 2.8}
mca2_vars_df = data.frame(mca2$colcoord, Variable = rep(names(cats), cats))
rownames(mca2_vars_df) = mca2$levelnames

# plot
ggplot(data = mca2_vars_df, 
       aes(x = X1, y = X2, label = rownames(mca2_vars_df))) +
  geom_hline(yintercept = 0, colour = "gray70") +
  geom_vline(xintercept = 0, colour = "gray70") +
  geom_text(aes(colour = Variable)) +
  labs(title = "MCA plot of variables using R package ca and ppglot2",
       subtitle = "Model: mjca(adj, lambda = \"Burt\", nd = 5)",
       caption = "Source: Kosheleva 2017")
```

### 2.9 Plot individuals and variables with ggplot2
Use the row coordinates `rowcoord` to make a data frame and pass it to ggplot():
```{r 2.9}
# individual observation coordinates 
mca2_obs_df = data.frame(mca2$rowcoord)
head(mca2_obs_df)

# variable coordinates  
mca2_vars_df = data.frame(mca2$colcoord, Variable = rep(names(cats), cats))
rownames(mca2_vars_df) = mca2$levelnames
head(mca2_vars_df)

# plot
mca1_obs_df = data.frame(mca1$ind$coord)

# plot of individuals and variables
ggplot(data = mca2_obs_df, aes(x = X1, y = X2)) +
  geom_hline(yintercept = 0, colour = "gray70") +
  geom_vline(xintercept = 0, colour = "gray70") +
  geom_point(aes(colour = factor(adj$Like)), alpha = 0.7) +
#  scale_colour_manual(values = c("black", "red")) + 
  geom_density2d(colour = "gray60", alpha = 0.5) +
  geom_text(data = mca2_vars_df, 
            aes(x = X1, y = X2, 
                label = rownames(mca2_vars_df), colour = Variable)) +
  scale_colour_discrete(name = "Variable") +
  labs(title = "MCA plot of variables using R package ca and ppglot2",
       subtitle = "Model: mjca(adj, lambda = \"Burt\", nd = 5)",
       caption = "Source: Kosheleva 2017")
```


