---
title: "Statistics for Hackers"
subtitle: "(after Jake VanderPlas 2016)"
date: 14.06.2018
font-family: 'Brill'
transition: none
output:
  slidy_presentation:
    df_print: kable
    footer: "Presentation link: https://goo.gl/Uf5BYc"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
```

```{r, include=FALSE}
library(tidyverse)
```

### 1. Types of statistics

Saying popularized by Mark Twain:

> There are three kinds of lies: lies, damned lies, and statistics.

* Frequentism
* Bayesianism
* Hackers' approach

### 2. Simulation

Whether a coin is fair: 27 heads and 13 tails

The probability of obtaining h heads in n tosses of a coin with a probability of heads equal to p is given by the binomial distribution:

$$P(H = h|p, n) = \binom{n}{h}\times p^h\times(1-p)^{1-h}$$

*Frequentist*: binomial test

* H$_0$ coin is fair
* α = 0.05

```{r, echo = FALSE}
h <- 1:40
data.frame(h = h, 
           p = dbinom(x = h, size = 40, prob = 0.5)) %>% 
  ggplot(aes(h, p))+
  geom_line()+
  geom_area(aes(x = ifelse(h>=qbinom(0.025, 40, 0.5) &
                           h<=qbinom(0.975, 40, 0.5), h, NA)), fill = "darkgreen")+
  geom_vline(xintercept = 27, linetype = 2)+
  scale_x_continuous(breaks = c(0:4*10, 27))+
  theme_bw()
```

So we can reject H$_0$ on p < 0.05.

*Hacker*: simulation

* Just simulate it!

```{r, cache= TRUE}
library(mosaic)
set.seed(42)
do(1000)*
  sum(sample(0:1, 40, replace = T)) ->
  simulations
```

```{r}
simulations %>% 
  mutate(greater = sum >= 27) %>% 
  group_by(greater) %>% 
  summarise(number = n())

simulations %>% 
  ggplot(aes(sum))+
  geom_density()+
  geom_vline(xintercept = 27, linetype = 2)+
  theme_bw()
```

### 3. Shuffling

```{r}
rus_est <- read.csv("https://goo.gl/11qut0")

rus_est %>% 
  group_by(language) %>% 
  summarise(mean = mean(speech_rate))

rus_est %>% 
  ggplot(aes(speech_rate, fill = language, color = language))+
  geom_density(alpha = 0.9)+
  geom_rug()+
  theme_bw()
```

Is this difference of 0.549683 statistically significant?

* mean speech rate for russians: 5.382458
* mean speech rate for estonians: 4.826735
* difference: 0.555723

*Frequentist*: Two-sample t-test

* H$_0$ Difference is not statistically significant.
* α = 0.05
* Welch's t-statistics

$$t = \frac{\mu_A - \mu_B}{\sqrt{\frac{var_A}{n_A}+\frac{var_B}{n_B}}}$$

$\mu$ -- mean of each group

$var$ -- variance estimation

$n$ -- number of observation in each group

```{r}
t <- t.test(rus_est[rus_est$language == "russian", ]$speech_rate,
       rus_est[rus_est$language == "estonian", ]$speech_rate)

t.value <- t$statistic
t.value
```


* find the degrees of freedom

$$d.f. = \frac{(var_A/n_A+var_B/n_B)^2}{\frac{var_A/n_A^2}{n_A-1}+\frac{var_B/n_B^2}{n_B-1}}$$

```{r}
t <- t.test(rus_est[rus_est$language == "russian", ]$speech_rate,
       rus_est[rus_est$language == "estonian", ]$speech_rate)

df <- t$parameter
df
```

```{r}
x <- seq(0, 5, 0.1)

data.frame(x, y = dt(x = x, df = df)) %>% 
  ggplot(aes(x, y))+
  geom_line()+
  geom_area(aes(x = ifelse(x>=qt(0.975, df), x, NA)), fill = "darkgreen")+
  geom_vline(xintercept = t.value, linetype = 2)+
  theme_bw()
```

So there is `r 1 - pt(t.value, df)` probability to see this or more extreme result giving H$_0$ is true.

*Hacker*: shuffling

If the language really don’t matter, then switching them shouldn’t change the result.

```{r, cache=TRUE}
set.seed(42)
do(1000) *
  (rus_est %>%
     mutate(speech_rate = shuffle(speech_rate)) %>%
     group_by(language) %>%
     summarize(mean_value = mean(speech_rate))) ->  
  many.shuffles
tail(many.shuffles)
```

Calculate the difference:
```{r}
many.shuffles %>% 
  group_by(.index) %>% 
  summarise(diff = diff(mean_value)) ->
  shuffle.diff
tail(shuffle.diff)

shuffle.diff %>% 
  mutate(greater = diff >= 0.555723) %>% 
  group_by(greater) %>% 
  summarise(number = n())

shuffle.diff %>% 
  ggplot(aes(x = diff)) +
    geom_density(fill = "darkgreen")+
    geom_rug()+
  geom_vline(xintercept = 0.555723, linetype = 2)+
  theme_bw()
```

### 4. bootstrapping

Calculate 95% confidence interval for mean [s] duration variable for heterosexual speakers from our orientation set.

95% CI formula again: $mean \pm 1.96\frac{standard\ deviation}{\sqrt{number\ of\ observation}}$

```{r}
homo <- read.csv("http://goo.gl/Zjr9aF")
homo %>% 
  group_by(orientation) %>%
  summarise(mean = mean(s.duration.ms),
            CI = 1.96*sd(s.duration.ms)/sqrt(length(s.duration.ms)))
```

Simulate the distribution by drawing samples with replacement.

```{r, cache=TRUE}
library(bootstrap)

data <- homo[homo$orientation == "homo", "s.duration.ms"]

set.seed(42)
boot_mean <- bootstrap(data, nboot = 1000, theta = mean)
boot_mean$thetastar %>% 
  data.frame(data = .) %>% 
  ggplot(aes(data))+
  geom_density(fill = "darkgreen")+
  geom_vline(xintercept = mean(data), linetype = 2)+
  theme_bw()

boot_mean$thetastar %>% 
  data.frame(data = .) %>% 
  summarise(mean = mean(data),
            CI = 1.96*sd(data)/sqrt(length(data)))
```


